---
title: "New mediator genes"
author: "Jinliang Yang"
date: "1-12-2021"
output: NULL
---


## Normalize the path:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
knitr::opts_knit$set(root.dir=normalizePath('../../')) 
```

#### New results

- Direct SNP: /common/jyanglab/zhikaiyang/projects/mediation/largedata/dsnps_vs_gwas/dsnps_34013rows.csv
- Indirect SNP: /common/jyanglab/zhikaiyang/projects/mediation/largedata/isnps_vs_feature/isnps_47143rows.csv
- Mediators: data/mediators_18812rows_by_cat.csv

### Old results for microbiome data

- cache/microbiome_mediators_34155rows.csv



### Mediators

```{r}
m <- read.csv("data/mediators_18812rows_by_cat.csv")
table(m$method)

### only consider std nitrogen condition
mi <- read.csv("cache/microbiome_mediators_34155rows.csv")

```
- eq: MedFix0.5
- fixed: MedFixBIC

- e2m: p-values for the exposure effect on the individual mediators
- m2y: p-values for the mediator effect on phenotype
- e2m2y: min value among c(1, max(e2m, m2y))

```{r}
# res.eq
m1 <- subset(m, method %in% "res.fixed" & padj <= 0.05)

m2 <- subset(mi, treatment %in% "stdN" & method %in% "res.fixed" & padj <= 0.05)

df <- rbind(m1[, c("id", "tissue", "trait")], m2[, c("id", "tissue", "trait")])
#write.table(df, "largedata/mediators_by_tissue.csv", sep=",", row.names=FALSE, quote=FALSE)

```

### merge two datasets together

```{r}


```

# Number of mediator genes by category and trait

```{r}
library(eulerr)
library(data.table)

# df$tissue <- factor(df$tissue, levels = c( "GRoot", "GShoot", "L3Base", "L3Tip", "LMAD", "LMAN","Kern"),
#                    labels =c( "Germinating Root", "Germinating Shoot", "Leaf 3 base", "Leaf 3 tip", "Adult leaf # during the day", "Adult leaf at night","Kernels after pollination"))


root = subset(df, tissue == "GRoot")$id
root <- root[!duplicated(root)]

shoot = subset(df, tissue == "GShoot")$id
shoot <- shoot[!duplicated(shoot)]
  
l3base = subset(df, tissue == "L3Base")$id
l3base <- l3base[!duplicated(l3base)]

l3tip = subset(df, tissue == "L3Tip")$id
l3tip <- l3tip[!duplicated(l3tip)]

kern = subset(df, tissue == "Kern")$id
kern <- kern[!duplicated(kern)]

df1 <- list(Root = root,
           Shoot = shoot,
           L3Base = l3base,
           L3Tip = l3tip,
           Kern = kern)

pdf("graphs/venn_mediators_by_tissue.pdf", width = 9, height= 6)
plot(venn(df1))
dev.off()


```


These results suggested that mediation effects were tissue specific.




# Number of mediator genes by category and trait


```{r}
library(eulerr)
library(data.table)


root = subset(m1, tissue == "GRoot")$id
root <- root[!duplicated(root)]

shoot = subset(m1, tissue == "GShoot")$id
shoot <- shoot[!duplicated(shoot)]
  
l3base = subset(m1, tissue == "L3Base")$id
l3base <- l3base[!duplicated(l3base)]

l3tip = subset(m1, tissue == "L3Tip")$id
l3tip <- l3tip[!duplicated(l3tip)]

lmad = subset(m1, tissue == "LMAD")$id
lmad = lmad[!duplicated(lmad)]

lman = subset(m1, tissue == "LMAN")$id
lman = lman[!duplicated(lman)]

kern = subset(m1, tissue == "Kern")$id
kern <- kern[!duplicated(kern)]

df1 <- list(Root = root,
           Shoot = shoot
          )

df2 <- list(L3Base = l3base,
           L3Tip = l3tip
           )
df3 = list(LMAD = lmad, LMAN = lman)

pdf("graphs/venn_mediators_by_root-shoot.pdf", width = 9, height= 6)
plot(venn(df1))
dev.off()

pdf("graphs/venn_mediators_by_l3base-l3tip.pdf", width = 9, height= 6)
plot(venn(df2))
dev.off()

pdf("graphs/venn_mediators_by_lmad-lman.pdf", width = 9, height= 6)
plot(venn(df3))
dev.off()


m1$tissue <- factor(m1$tissue, levels = c( "GRoot", "GShoot", "L3Base", "L3Tip", "LMAD", "LMAN","Kern"),
                    labels =c( "Germinating Root/Shoot", "Germinating Root/Shoot", "Leaf 3 base/tip", "Leaf 3 base/tip", "Adult leaf during the day", "Adult leaf at night","Kernels after pollination"))

shoot_root = subset(m1, tissue == "Germinating Root/Shoot")$id
shoot_root <- shoot_root[!duplicated(shoot_root)]

  
l3base_tip = subset(m1, tissue == "Leaf 3 base/tip")$id
l3base_tip <- l3base_tip[!duplicated(l3base_tip)]


lmad = subset(m1, tissue == "Adult leaf during the day")$id
lmad = lmad[!duplicated(lmad)]

lman = subset(m1, tissue == "Adult leaf at night")$id
lman = lman[!duplicated(lman)]

kern = subset(m1, tissue == "Kernels after pollination")$id
kern <- kern[!duplicated(kern)]


df5 = list(Shoot_Root = shoot_root, L3Base_Tip = l3base_tip, LMAD = lmad, LMAN = lman, Kern = kern)

pdf("graphs/venn_mediators_by_7tissues.pdf", width = 9, height= 6)
plot(venn(df5))
dev.off()

```


