---
title: "Mediator genes for all tissues and traits"
author: "Jinliang Yang"
date: "1-21-2021"
output: NULL
---


## Normalize the path:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
knitr::opts_knit$set(root.dir=normalizePath('../../')) 
```

#### New results

- Direct SNP: /common/jyanglab/zhikaiyang/projects/mediation/largedata/dsnps_vs_gwas/dsnps_34013rows.csv
- Indirect SNP: /common/jyanglab/zhikaiyang/projects/mediation/largedata/isnps_vs_feature/isnps_47143rows.csv
- Mediators: data/mediators_18812rows_by_cat.csv

### Old results for microbiome data

- cache/microbiome_mediators_34155rows.csv



### Mediators


```{r}
m <- read.csv("data/mediators_18812rows_by_cat.csv")
table(m$method)

### only consider std nitrogen condition
mi <- read.csv("cache/microbiome_mediators_34155rows.csv")

# res.eq
m1 <- subset(m, method %in% "res.fixed" & padj <= 0.05)
m1$type <- "Morphology"
m1[m1$category == "Metabolite", ]$type <- "Metabolite"


m2 <- subset(mi, treatment %in% "stdN" & method %in% "res.fixed" & padj <= 0.05)
m2$type <- "Microbiome"

m1 <- m1[, c("type", "category", "tissue", "trait", "id", "method", "e2m", "m2y","e2m2y","padj","coef")]
m2 <- m2[, c("type", "treatment", "tissue", "trait", "id", "method", "e2m", "m2y","e2m2y","padj","coef")]
names(m2)[2] <- "category"

df <- rbind(m1, m2) 
```
- eq: MedFix0.5
- fixed: MedFixBIC

- e2m: p-values for the exposure effect on the individual mediators
- m2y: p-values for the mediator effect on phenotype
- e2m2y: min value among c(1, max(e2m, m2y))



# Polygenic mediators


V3 => V4 gene id and coordinates:

```{r}
V3_V4 <- fread("/common/jyanglab/zhikaiyang/projects/mediation/largedata/V3_V4.csv", 
               header = FALSE, data.table=FALSE)
V3_V4 <- V3_V4[,c(1,2)]
Zea_mays.AGPv4_anno <- fread("/common/jyanglab/zhikaiyang/projects/mediation/largedata/Zea_mays.AGPv4_anno.txt",
                             header = TRUE, data.table = FALSE)
v34 <- merge(V3_V4, Zea_mays.AGPv4_anno, by.x = "V2", by.y = "ensembl_gene_id_v4")
fwrite(v34, "cache/gene_v34_conversion.txt", sep="\t", row.names=FALSE, quote=FALSE)
```


Find the top mediators

```{r}
library(plyr)
library("data.table")

df <- m1
pall <- ddply(df, .(id), nrow)
pall <- pall[order(pall$V1, decreasing = T),]

v34 <- fread("cache/gene_v34_conversion.txt", data.table=FALSE)

p2 <- merge(pall, v34[, 1:5], by.x="id", by.y="V1")
p2 <- p2[order(p2$V1, decreasing = T),]
```


```{r}
library(huskeR)

names(p2)[4:5] <- c("chr", "pos")
#mediator genes ploted in manhattan plot
out <- plot_mht(inputdf=p2, hdf=NULL, col2plot = "V1", jmph = FALSE, pl = TRUE,
  cl_file = "data/chr_length_B73v4.csv", cex = 1.5,
  pch = 16, col = rep(c("slateblue", "cyan4"), 5), GAP = 5e+06,
  yaxis = NULL, ylab="Number of mediated traits")

pdf("graphs/mht_num_med_traits.pdf", width = 15, height= 5)
plot_mht(inputdf=p2, hdf=NULL, col2plot = "V1", jmph = FALSE, pl = TRUE,
  cl_file = "data/chr_length_B73v4.csv", cex = 1.5,
  pch = 16, col = rep(c("slateblue", "cyan4"), 5), GAP = 5e+06,
  yaxis = NULL, ylab="Number of mediated traits")
dev.off()





```




```{r}
#library(qqman)
#p2new = data.frame(SNP = p2$V2, CHR = as.numeric(as.character(p2$chr)), BP = p2$pos, P = p2$V1)

#manhattan(subset(p2new, CHR == 2), highlight = c("Zm00001d00771"), xlim = c(238043557,	238245188), main = "Chr 2", ylim= c(0, 40), type ="h")

p2new = subset(p2, chr == "2")
x = rep(0,max(p2new$end_pos))
x <- runif(15000)
x[p2new$pos] <- p2new$V1

plot(x, type="h", col= "green", xlim = c(238043557,	238245188))
rm(x)
```




