---
title: "Post prunning statistics"
output: NULL
date: 04-22-2020
author: "Zhikai Yang & Jinliang Yang"
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE) 
knitr::opts_knit$set(root.dir=normalizePath('../../')) 
#library(tidyverse)
```

# Analysis of LD decay rate

After SNP prunning, calculate some basic stats, i.e MAF, missingness, distribution, LD decay, PCs

```{bash, eval=FALSE}
#calculate MAF, Missingness
module load plink

plink -bfile allchr_bisnp_n282_snpid_maf01_geno2_pruned --freq --missing --out allchr_bisnp_n282_snpid_maf01_geno2_pruned


```

```{r}
#should be done in high performance computing environment

## Plot MAF and Missingness

library("data.table")

maf <- fread("largedata/geno_282/allchr_bisnp_n282_snpid_maf01_geno2_pruned.frq", header=TRUE)
lmiss <- fread("largedata/geno_282/allchr_bisnp_n282_snpid_maf01_geno2_pruned.lmiss", header=TRUE)


pdf("graphs/maf_lmiss_pruned.pdf", width = 10, height=5)
par(mfrow=c(1,2))
hist(maf$MAF, breaks=100, col="#cdb79e", main="MAF (SNP = 696548)", xlab="Minor Allele Freq")
abline(v=0.01, lty=2, col="black", lwd=3)
abline(v=0.05, lty=2, col="red", lwd=3)

hist(lmiss$F_MISS, breaks=100, col="#cdb79e", main="Missingness (SNP = 696548)", xlab="Missing Rate")
abline(v=0.6, lty=2, col="red", lwd=3)
abline(v=0.2, lty=2, col="red", lwd=3)
dev.off()

```
----
```{bash, eval=FALSE}
# LD decay
plink --bfile allchr_bisnp_n282_snpid_maf01_geno2_pruned  --ld-window 10 --ld-window-kb 100 --r2 --ld-window-r2 0  --out allchr_bisnp_n282_snpid_maf01_geno2_pruned
```

```{r}
#should be done in high performance computing environment

## Plot LD decay

library(data.table)
ld <- fread("largedata/geno_282/allchr_bisnp_n282_snpid_maf01_geno2_pruned.ld", data.table=FALSE)
ld$dis <- ld$BP_B - ld$BP_A
# summarize for each bp distance, the mean R2 
ld <- as.data.table(ld)

df <- ld[,.(meanr2=mean(R2), sd=sd(R2)), by=dis]
write.table(df, "cache/ld_mean_sd_pruned.csv", sep=",", quote=FALSE)

df <- read.csv("cache/ld_mean_sd_pruned.csv")
library(ggplot2)

pdf("graphs/LD_decay_pruned.pdf", width = 10, height=5)
par(mfrow=c(1,2))
p1 <- ggplot(df, aes(x=dis, y=meanr2)) +
    #facet_grid(. ~ type) +
    geom_point(size=2) +
    geom_smooth(method='gam') +
    theme_bw() +
    xlab("physitcal distance") +
    ylab("LD (r2)") +
    theme(axis.text=element_text(size=12), axis.title=element_text(size=15) )
  #########
p1

df <- as.data.frame(df)
df <- subset(df, dis < 5000)
p2 <- ggplot(df, aes(x=dis, y=meanr2)) +
    #facet_grid(. ~ type) +
    geom_point(size=2) +
    geom_smooth(method='loess') +
    theme_bw() +
    xlab("physitcal distance") +
    ylab("LD (r2)") +
    theme(axis.text=element_text(size=12), axis.title=element_text(size=15) )
  #########
p2
dev.off()
```
### PCs

--pca [count] ['header'] ['tabs'] ['var-wts']

--pca-cluster-names <name(s)...>
--pca-clusters <filename>

By default, --pca extracts the top 20 principal components of the variance-standardized relationship matrix; you can change the number by passing a numeric parameter. Eigenvectors are written to plink.eigenvec, and top eigenvalues are written to plink.eigenval. The 'header' modifier adds a header line to the .eigenvec file(s), and the 'tabs' modifier makes the .eigenvec file(s) tab- instead of space-delimited. You can request variant weights with the 'var-wts' modifier, and dump the matrix by using --pca in combination with --make-rel/--make-grm-gz/--make-grm-bin.

```{bash, eval=FALSE}
cd largedata/geno_282

plink -bfile allchr_bisnp_n282_snpid_maf01_geno2_pruned --pca 'tabs' --out allchr_bisnp_n282_snpid_maf01_geno2_pruned
```


## Plot results of the PCA

### Plot 3D PCA

```{r}
#should be done in high performance computing environment

library(ggplot2)
library(cowplot)
library(plyr)

df <- read.delim("largedata/geno_282/allchr_bisnp_n282_snpid_maf01_geno2_pruned.eigenvec", header=FALSE)
write.table(df, "cache/allchr_bisnp_n282_snpid_maf01_geno2_pruned.eigenvec", se="\t", row.names=FALSE, quote=FALSE)

df <- read.delim("cache/allchr_bisnp_n282_snpid_maf01_geno2_pruned.eigenvec", header=TRUE)
names(df)[3:5] <- c("PC1", "PC2", "PC3")



#install.packages("scatterplot3d")
library("scatterplot3d") # load

fsize=16

pdf("graphs/Fig_pca1-3_maize.pdf", width=10, height=10)
scatterplot3d(df[,3:5], pch = 16, cex.symbol=1.2, color="#00BFC4", main="Maize Diversity Panel", angle=40)
dev.off()

```




