---
title: "Prepare data for Med analysis: learned from Zhikai"
author: "Jinliang Yang"
date: "09-23-2020"
output: NULL
---


## Normalize the path:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
knitr::opts_knit$set(root.dir=normalizePath('../../')) 
```

### RNA-seq data


```{r}
rna_filtering <- function(rna, cutoff_var = 0.05, cutoff_mean = 1){
  # The purpose of this function is to split the RNA-seq data into different parts by tissue type.
  # And then filter the variance <= cutoff_var and mean <= cutoff_mean!
  
  # rna: data.frame of the RNA-seq read count. [data.frame, object]
  # cutoff_var: cutoff for variances among the samples. [num, = 0.05] 
  # cutoff_mean: cutoff for means among the genotypes. [num, = 1]
  
  tissue <- table(rna$TissueWODate)
  tissue <- tissue[tissue > 100]
  ts <- data.frame(tissue)
  
  for(i in 1:length(tissue)){
    gs <- subset(rna, TissueWODate %in% ts$Var1[i])
    #dim(gs) # 295 37136
    message(sprintf("###>>> working on [ %s trait] with dim [ row: %s, col: %s ]", ts$Var1[i], nrow(gs), ncol(gs)))
    # remove the duplicated ones
    gs0 <- gs[!duplicated(gs$HMP32Name), ]
    message(sprintf("###>>> after removing duplicated records, [ row: %s, col: %s ] remaning", nrow(gs0), ncol(gs0)))
    #dim(gs0) #278 37136

    #determine the first column of the gene expression
    idx1 <- which(names(gs0) == "AC148152.3_FG001")

    M <- gs0[, c(6, idx1:ncol(gs0))]
    M$HMP32Name <- gsub("282set_", "", M$HMP32Name)

    variances <- apply(M[, 2:ncol(M)], 2, var)
    M <- M[, -(which(variances <= 0.05) + 1)] # 278 25038
    message(sprintf("###>>> after removing [variances < %s], [ row: %s, col: %s ] remaning", cutoff_var, nrow(M), ncol(M)-1))
   
    means <- apply(M[,2:ncol(M)], 2, mean)
    M <- M[, -(which(means <=1 ) + 1)] #278 19357
    message(sprintf("###>>> after removing [mean < %s], [ row: %s, col: %s ] remaning", cutoff_mean, nrow(M), ncol(M)-1))
    
    outfile <- paste0("largedata/RNA-seq/filtered_", ts$Var1[i], ".csv")
    message(sprintf("###>>> output results to [ %s ]", outfile))
    message(sprintf("\n"))
    names(M)[1] <- "genotype"
    fwrite(M, outfile, sep=",", row.names=FALSE, quote=FALSE)
  }
}
```


```{r}
library(data.table)

rna <- fread("/common/jyanglab/shared/dbcenter/Kremling_Nature3RNASeq282_March2018/Expression_matrix/df_STAR_HTSeq_counts_B73_match_based_on_genet_dist_DESeq2_normed_fpm_rounded_origNames_and_Altnames.txt", data.table=FALSE)

rna_filtering(rna, cutoff_var = 0.05, cutoff_mean = 1)
  
```

###>>> working on [ GRoot trait] with dim [ row: 291, col: 37136 ]
###>>> after removing duplicated records, [ row: 273, col: 37136 ] remaning
###>>> after removing [variances < 0.05], [ row: 273, col: 25600 ] remaning
###>>> after removing [mean < 1], [ row: 273, col: 19206 ] remaning
###>>> output results to [ largedata/RNA-seq/filtered_GRoot.csv ]


###>>> working on [ GShoot trait] with dim [ row: 295, col: 37136 ]
###>>> after removing duplicated records, [ row: 278, col: 37136 ] remaning
###>>> after removing [variances < 0.05], [ row: 278, col: 25037 ] remaning
###>>> after removing [mean < 1], [ row: 278, col: 19356 ] remaning
###>>> output results to [ largedata/RNA-seq/filtered_GShoot.csv ]


###>>> working on [ Kern trait] with dim [ row: 254, col: 37136 ]
###>>> after removing duplicated records, [ row: 229, col: 37136 ] remaning
###>>> after removing [variances < 0.05], [ row: 229, col: 24579 ] remaning
###>>> after removing [mean < 1], [ row: 229, col: 18226 ] remaning
###>>> output results to [ largedata/RNA-seq/filtered_Kern.csv ]


###>>> working on [ L3Base trait] with dim [ row: 302, col: 37136 ]
###>>> after removing duplicated records, [ row: 263, col: 37136 ] remaning
###>>> after removing [variances < 0.05], [ row: 263, col: 25048 ] remaning
###>>> after removing [mean < 1], [ row: 263, col: 19108 ] remaning
###>>> output results to [ largedata/RNA-seq/filtered_L3Base.csv ]


###>>> working on [ L3Tip trait] with dim [ row: 295, col: 37136 ]
###>>> after removing duplicated records, [ row: 265, col: 37136 ] remaning
###>>> after removing [variances < 0.05], [ row: 265, col: 26764 ] remaning
###>>> after removing [mean < 1], [ row: 265, col: 18546 ] remaning
###>>> output results to [ largedata/RNA-seq/filtered_L3Tip.csv ]


###>>> working on [ LMAD trait] with dim [ row: 210, col: 37136 ]
###>>> after removing duplicated records, [ row: 204, col: 37136 ] remaning
###>>> after removing [variances < 0.05], [ row: 204, col: 25804 ] remaning
###>>> after removing [mean < 1], [ row: 204, col: 18304 ] remaning
###>>> output results to [ largedata/RNA-seq/filtered_LMAD.csv ]


###>>> working on [ LMAN trait] with dim [ row: 276, col: 37136 ]
###>>> after removing duplicated records, [ row: 260, col: 37136 ] remaning
###>>> after removing [variances < 0.05], [ row: 260, col: 25630 ] remaning
###>>> after removing [mean < 1], [ row: 260, col: 18453 ] remaning
###>>> output results to [ largedata/RNA-seq/filtered_LMAN.csv ]







